---
title: "ST 558 Project 2"
author: "Holly Probasco"
format: pdf
editor: visual
---

## Libraries
```{r}
library(tidyverse)
library(ggplot2)
```


## Accessing the API
```{r}
query_function <- function(subject, set.name = "", pokemon.name = "", sort.by = "number", select = NULL) {
  if(set.name == ""){ 
    data_url = paste0("https://api.pokemontcg.io/v2/", subject, "?q=name:", pokemon.name, "&orderBy:", sort.by, select) 
  } else if(pokemon.name == "") {
    data_url = paste0("https://api.pokemontcg.io/v2/", subject, "?q=set.name:", set.name, "&orderBy:", sort.by, select)
  } else {
  data_url = paste0("https://api.pokemontcg.io/v2/", subject, "?q=set.name:", set.name,"%20name:", pokemon.name, "&orderBy:", sort.by, select) }
info <- httr::GET(url=data_url)
parsed = jsonlite::fromJSON(rawToChar(info$content))
queried_data = as.tibble(parsed$data)
return(queried_data)
}
```


### Example using the function
This will return all Mew cards from the Celebrations set, sorted by card number
```{r}
query_function(subject = "cards", set.name = "celebrations", pokemon.name = "mew", sort.by = "number")
```


### another type of data able to be accessed 
This will return all the different types of cards, sorted by name
```{r}
query_function(subject = "types", sort.by = "name")
```

### Using select
This will return all Pikachu cards in the Base set, specifically calling for the inclusion of the column of card rarity.
```{r}
raritycall = query_function("cards", set.name = "Base", pokemon.name = "Pikachu", select = c("rarity"))

#calling the selected column to show
raritycall$rarity
```


## Know How to Summarize the Data
### You should create some contingency tables
This is a summary of the Yveltal cards in the set XY based on artist
```{r}
contingency_data = query_function("cards", set.name = "XY", pokemon.name = "Yveltal")
cont_table = table(contingency_data$name, contingency_data$artist)
cont_table
```
Based on this table, we can see that 5ban Graphics did most of the artwork for this set's Yveltal cards, doing 5 out of 8.

Doing another table, we can see the breakdown of what type of card this artwork was used on
```{r}
cont_table2 = table(contingency_data$name,contingency_data$set$name)
cont_table2
```
Looks like the 5ban Graphics artwork was used on the promotional cards for this set, while the other 3 cards were used in the main set itself.


Let's look at one more contingency table for something different, the fast attack for each Yveltal card 
```{r}
# first break down the attacks column to only include the first attack
contingency_attacks = sapply(contingency_data$attacks, function(x) {
  if (!is.null(x)) { x$name[1] } else { NA }
}) # use sapply so that just the first row is taken from each row in the attacks data frame
cont_table = table(contingency_data$name, contingency_attacks)
cont_table
```

### create numerical summaries for some quantitative variables at each setting of some of your categorical variables
we can do this using the summarize function. Lets look through every released Charizard card across all sets, and compare how much Health Points they all have
```{r}
Char_data <- as.data.frame(query_function("cards", pokemon.name = "Charizard"))
numsum_data <- 
  Char_data |> 
  group_by(setname = set$name) |> #group by set to take the mean of all cards in one set
    summarize(cardcount = n(), meanhp = mean(as.numeric(hp),na.rm=TRUE), medianhp = median(as.numeric(hp),na.rm=TRUE), sdhp = sd(as.numeric(hp), na.rm = TRUE), IQRhp = IQR(as.numeric(hp), na.rm = TRUE))
numsum_data |> arrange(desc(numsum_data$meanhp))
```
This shows that the largest meanhp for Charizard is 330 in 4 different sets, whereas the smallest is only 80 hp in the Team Rocket set. This is an older set, so this makes sense.

### Create at least four plots
#### Plot 1 
shows all Charizard cards by set, with mean hp over 200, as well as only including sets with more than 1 card
```{r}
plot1_data <- numsum_data |> filter(cardcount > 1 & meanhp > 200) 
ggplot(plot1_data, aes(x = setname, y = cardcount)) + geom_col(fill = "red") + coord_flip() + labs(x = "Set Name", x = "Card Count", title = "Charizard popularity in sets")
```
From this graph it looks like Charizard is popular enough to be consistently included in sets that are released.


#### Plot 2
This plot is of the damage counters of the Fast attack of each Charizard card, faceted by type
```{r}
# for easier plotting, create smaller data frame with just the needed information
type <- sapply(Char_data$types, function(x) {
  if (!is.null(x) && length(x) < 2) { x[1] } else { NA } 
})

chplotdata = data.frame(Pokemon = Char_data$name, type = type, hp = Char_data$hp, rarity = Char_data$rarity, cost = Char_data$tcgplayer$prices$holofoil$high)

chplot <- ggplot(chplotdata |> drop_na(type, cost) , aes(x = cost, y = hp, color = type))

chplot + geom_point() +
geom_smooth(method = lm) +
theme_light()
```


